The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\ with the Owens Valley Long Wavelength Array

### 2017-09-05 paper.tex
commit channel maps and latex updates

Studies of the cosmic microwave background (CMB) have given us an unprecedented understanding of the
universe during the epoch of recombination \citep{2013ApJS..208...19H, 2014A&A...571A..16P,
2016A&A...594A..13P}. Before arriving at the Earth, the light from the CMB propagates through the
intergalactic medium (IGM), intervening galaxies and galaxy clusters, and the Milky Way's
interstellar medium (ISM). As photons from the CMB propagate through this intervening material,
scattering, absorption, or additional emission can perturb or corrupt the thermal CMB spectrum. In
order to measure the pristine CMB, these foreground components (eg. synchrotron emission, free-free
emission, thermal dust emission) are a nuisance and careful modeling and removal of these components
is essential \citep{2016A&A...594A..10P, 2016A&A...594A..25P}.

However in many cases the study of CMB foregrounds is scientifically interesting in its own right.
Measurements of dust-correlated anomalous microwave emission \citep{1997ApJ...486L..23L} led to the
discovery of spinning dust emission \citep{1998ApJ...508..157D}. Thomson scattering in the IGM
attenuates the CMB spectrum, but the optical depth implies a mean redshift to the Epoch of
Reionization (EoR) of $z_r = 8.8^{+1.7}_{-1.4}$ \citep{2016A&A...594A..13P}.  The thermal
Sunyaev-Zel'dovich effect \citep{1970CoASP...2...66S, 1972CoASP...4..173S} has been used to discover
new galaxy clusters \citep{2010ApJ...722.1180V} and measure the distance to galaxy clusters when
combined with X-ray measurements \citep{2006ApJ...647...25B}. The kinetic Sunyaev-Zel'dovich effect
has been used to constrain the duration of reionization $\Delta z_r \le 4.4$
\citep{2012ApJ...756...65Z}.

@@

\begin{figure}[t]
    \plotone{figures/lcurve/lcurve}
    \caption{
        An example L-curve using OVRO-LWA data at 36.528 MHz. The $x$-axis is the norm of the
        solution (in this case the spherical harmonic coefficients) given in arbitrary units. The
        $y$-axis is the least-squares norm given in arbitrary units. $\varepsilon$ is the
        regularization parameter. When the regularization parameter is small, the norm of the
        solution grows rapidly. When the regularization parameter is large, the least-squares norm
        grows rapidly.
    }
    \label{fig:lcurve}
\end{figure}

@@

Altogether the adaptation of CLEAN applied to the maps presented in this paper is summarized below.
\begin{algorithmic}[1]
    \Require{$\b a$ is the solution to Equation~\ref{eq:tikhonov-solution}}
    \Function{CLEAN}{$\b a$}
    \Let{$\b M$}{$\b B^*\b B^*$}
    \Let{$\b U$}{${\rm chol}(\b B)$} \Comment{Cholesky decomposition}
    \While{noise in map $>$ threshold}
    \State find $N$ pixels with the largest residual flux
    \Let{$\b x$}{$\sum_{i=1}^N \,(\text{pixel flux}) \times \b a_\text{PS}(\theta_i, \phi_i)$}
    \Let{$\b y$}{$\b U^{-1}\big(\b U^*\big)^{-1}\b M\b x$}
    \Let{$\b a$}{$\b a - (\text{loop gain})\times\b y$}
    \State record subtracted components
    \EndWhile
    \Let{$\b a$}{$\b a + (\text{restored components})$}
    \State \Return{$\b a$}
    \EndFunction
\end{algorithmic}

@@

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/psf/psf}
    \caption{
        \textbf{(left panel)}
        This figure shows the antenna layout for the OVRO-LWA. Black dots correspond to antennas
        within the 200 m diameter core of the array. The 32 open circles are the expansion antennas
        built in early 2016 in order to increase the longest baseline to 1.5 km. The 5 crosses are
        antennas equipped with noise-switched front ends.
        \textbf{(middle panel)}
        The computed $m$-mode analysis imaging PSF for the OVRO-LWA for a point source at a
        declination of $+45^\circ$ and frequency of $36.528$ MHz.
        \textbf{(right panel)}
        The computed $m$-mode analysis imaging PSF for the OVRO-LWA for a point source at a
        declination of $+45^\circ$ and frequency of $73.152$ MHz.
    }
    \label{fig:psf}
\end{figure*}

@@

\subsection{The Rainy Day Dataset}

The dataset used in this paper spans 28 consecutive hours beginning at 2017-02-17 12:00:00 UTC time.
This time was chosen based on the fact that it was raining at OVRO and rain tends to improve the
low-frequency RFI environment considerably. During this time the OVRO-LWA operated as a
zenith-pointing drift scanning interferometer.  The correlator dump time was selected to be 13
seconds such that the correlator output evenly divides a sidereal day.

The geomagnetic conditions during this time were mild. The Disturbance storm time (Dst) index was
$>-30$ nT during the entirety of the observing period.\footnote{
    The Dst index was obtained from the World Data Center for Geomagnetism, Kyoto University
    (\url{http://swdcwww.kugi.kyoto-u.ac.jp/}).
}
Following the classification scheme of \citet{2008GMS...181.....K}, a weak geomagnetic storm has
$\text{Dst} < -30$ nT. Therefore there were no geomagnetic storms during the time of these
observations.

@@

\begin{figure*}[t]
    \plottwo{figures/beam/source-tracks}{figures/beam/stokes-I-beam}
    \caption{
        (left panel) The tracks traced through the sky by the 8 brightest point sources visible to the
        OVRO-LWA.
        (right panel) The amplitude of the OVRO-LWA beam at 36.528 MHz as a result of the beam model
        fit using Zernike polynomials and the symmetry assumptions outlined in the text.
    }
    \label{fig:beam}
\end{figure*}

@@

@@

\begin{table*}[t]
    \centering
    \begin{tabular}{cccccccccc}
        \hline
        \hline
        & \tbf{Frequency} & \tbf{Bandwidth} & \tbf{FWHM} &
            \multicolumn2c{\tbf{Thermal Noise}} &&
            \multicolumn2c{\tbf{Spherical Harmonics}} &
            \tbf{Singular Values} \\
        \cline{5-6}
        \cline{8-9}
        \tbf{\#} & MHz & MHz & arcmin &
            K & Jy &&
            $l=0$ & $m=0$ &
            fraction $>\sqrt{\varepsilon}$ \\
        \hline
        1 & 36.528 & 0.024 & 16.2 & 372.8 & & & no & yes \\
        2 & 41.760 & 0.024 & 14.7 & 321.2 & & & no & yes \\
        3 & 46.992 & 0.024 & 14.2 & 247.0 & & & no & no \\
        4 & 52.224 & 0.024 & 13.6 & 205.7 & & & no & no \\
        5 & 57.456 & 0.024 & 13.5 & 167.7 & & & no & no \\
        6 & 62.688 & 0.024 & 13.5 & 137.6 & & & no & no \\
        7 & 67.920 & 0.024 & 13.5 & 114.9 & & & no & no \\
        8 & 73.152 & 0.024 & 13.7 &  93.4 & & & no & no \\
        \hline \hline
    \end{tabular}
    \caption{Summary of the generated all-sky maps\label{tab:summary}}
\end{table*}

@@

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/channel-maps/channel-maps}
    \caption{
        Mollweide-projected maps of the VHF sky north of $-30^\circ$ generated using the OVRO-LWA
        and $m$-mode analysis imaging. The properties of each map are summarized in
        Table~\ref{tab:summary}. The color scale is logarithmic and has units of Kelvin. As
        described in the text, 8 bright sources have been subtracted. The small blank region near
        the bottom of each map corresponds to the region of sky within 2 degrees of the Sun at the
        time of the observations.
    }
    \label{fig:channel-maps}
\end{figure*}

\begin{figure*}[p]
    \gridline{
        \fig{figures/jackknife-maps/4.pdf}
            {0.48\textwidth}{}
        \fig{figures/jackknife-maps/6.pdf}
            {0.48\textwidth}{}
    }
    \gridline{
        \fig{figures/jackknife-maps/8.pdf}
            {0.48\textwidth}{}
        \fig{figures/jackknife-maps/10.pdf}
            {0.48\textwidth}{}
    }
    \gridline{
        \fig{figures/jackknife-maps/12.pdf}
            {0.48\textwidth}{}
        \fig{figures/jackknife-maps/14.pdf}
            {0.48\textwidth}{}
    }
    \gridline{
        \fig{figures/jackknife-maps/16.pdf}
            {0.48\textwidth}{}
        \fig{figures/jackknife-maps/18.pdf}
            {0.48\textwidth}{}
    }
    \caption{
        The jackknife estimated standard deviation corresponding to the maps presented in
        Figure~\ref{fig:channel-maps}. The color scale is logarithmic and has units of Kelvin.
    }
    \label{fig:jackknife-maps}
\end{figure*}

>>>>

\todo{this entire section needs a rewrite, it is way too rambling}

%Studies of the cosmic microwave background (CMB) have given us an unprecedented understanding of the
%universe during the epoch of recombination \citep{2013ApJS..208...19H, 2014A&A...571A..16P,
%2016A&A...594A..13P}. Before arriving at the Earth, the light from the CMB propagates through the
%intergalactic medium (IGM), intervening galaxies and galaxy clusters, and the Milky Way's
%interstellar medium (ISM). As photons from the CMB propagate through this intervening material,
%scattering, absorption, or additional emission can perturb or corrupt the thermal CMB spectrum. In
%order to measure the pristine CMB, these foreground components (eg. synchrotron emission, free-free
%emission, thermal dust emission) are a nuisance and careful modeling and removal of these components
%is essential \citep{2016A&A...594A..10P, 2016A&A...594A..25P}.
%
%However in many cases the study of CMB foregrounds is scientifically interesting in its own right.
%Measurements of dust-correlated anomalous microwave emission \citep{1997ApJ...486L..23L} led to the
%discovery of spinning dust emission \citep{1998ApJ...508..157D}. Thomson scattering in the IGM
%attenuates the CMB spectrum, but the optical depth implies a mean redshift to the Epoch of
%Reionization (EoR) of $z_r = 8.8^{+1.7}_{-1.4}$ \citep{2016A&A...594A..13P}.  The thermal
%Sunyaev-Zel'dovich effect \citep{1970CoASP...2...66S, 1972CoASP...4..173S} has been used to discover
%new galaxy clusters \citep{2010ApJ...722.1180V} and measure the distance to galaxy clusters when
%combined with X-ray measurements \citep{2006ApJ...647...25B}. The kinetic Sunyaev-Zel'dovich effect
%has been used to constrain the duration of reionization $\Delta z_r \le 4.4$
%\citep{2012ApJ...756...65Z}.

@@

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/lcurve/lcurve}
    \caption{
        An example L-curve using OVRO-LWA data at 36.528 MHz. The $x$-axis is the norm of the
        solution (in this case the spherical harmonic coefficients) given in arbitrary units. The
        $y$-axis is the least-squares norm given in arbitrary units. $\varepsilon$ is the
        regularization parameter. When the regularization parameter is small, the norm of the
        solution grows rapidly. When the regularization parameter is large, the least-squares norm
        grows rapidly.
    }
    \label{fig:lcurve}
\end{figure}

@@

Altogether the adaptation of CLEAN applied to the maps presented in this paper is summarized below.
\begin{algorithmic}[1]
    \Require{$\b a$ is the solution to Equation~\ref{eq:tikhonov-solution}}
    \Function{CLEAN}{$\b a$}
    \Let{$\b M$}{$\b B^*\b B$}
    \Let{$\b U$}{${\rm chol}(\b M + \varepsilon\b I)$} \Comment{Cholesky decomposition}
    \While{noise in map $>$ threshold}
    \State find $N$ pixels with the largest residual flux
    \Let{$\b x$}{$\sum_{i=1}^N \,(\text{pixel flux}) \times \b a_\text{PS}(\theta_i, \phi_i)$}
    \Let{$\b y$}{$\b U^{-1}\big(\b U^*\big)^{-1}\b M\b x$}
    \Let{$\b a$}{$\b a - (\text{loop gain})\times\b y$}
    \State record subtracted components
    \EndWhile
    \Let{$\b a$}{$\b a + (\text{restored components})$}
    \State \Return{$\b a$}
    \EndFunction
\end{algorithmic}

@@

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/antenna-layout/antenna-layout}
    \caption{
        This figure shows the antenna layout for the OVRO-LWA. Black dots correspond to antennas
        within the 200 m diameter core of the array. The 32 open circles are the expansion antennas
        built in early 2016 in order to increase the longest baseline to 1.5 km. The 5 crosses are
        antennas equipped with noise-switched front ends.
    }
    \label{fig:antenna-layout}
\end{figure}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/psf/psf}
    \caption{
        \textit{(left panel)}
        This figure shows the antenna layout for the OVRO-LWA. Black dots correspond to antennas
        within the 200 m diameter core of the array. The 32 open circles are the expansion antennas
        built in early 2016 in order to increase the longest baseline to 1.5 km. The 5 crosses are
        antennas equipped with noise-switched front ends.
        \textit{(middle panel)}
        The computed $m$-mode analysis imaging PSF for the OVRO-LWA for a point source at a
        declination of $+45^\circ$ and frequency of $36.528$ MHz.
        \textit{(right panel)}
        The computed $m$-mode analysis imaging PSF for the OVRO-LWA for a point source at a
        declination of $+45^\circ$ and frequency of $73.152$ MHz.
    }
    \label{fig:psf}
\end{figure*}

@@

The dataset used in this paper spans 28 consecutive hours beginning at 2017-02-17 12:00:00 UTC time.
During this time the OVRO-LWA operated as a zenith-pointing drift scanning interferometer.  The
correlator dump time was selected to be 13 seconds such that the correlator output evenly divides a
sidereal day.

@@

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/beam/beam}
    \caption{
        \textit{(left panel)}
        The tracks traced through the sky by the 8 brightest point sources visible to the
        OVRO-LWA.
        \textit{(middle panel)}
        The amplitude of the OVRO-LWA beam at 36.528 MHz as a result of the beam model
        fit using Zernike polynomials and the symmetry assumptions outlined in the text.
        \textit{(right panel)}
        The amplitude of the OVRO-LWA beam at 73.152 MHz as a result of the beam model
        fit using Zernike polynomials and the symmetry assumptions outlined in the text.
    }
    \label{fig:beam}
\end{figure*}

@@

\subsection{Ionospheric Conditions}

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/vtec/vtec}
    \caption{
        The median vertical TEC within 200 km of OVRO during the time of the observation. The gray
        shaded regions indicate times outside of the observing period. The gray vertical lines
        indicate sunrise and sunset (as labeled).
    }
    \label{fig:vtec}
\end{figure}

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/scintillation-refraction/scintillation-refraction}
    \caption{
        Light curves for Cyg A and Cas A.
    }
    \label{fig:scintillation}
\end{figure*}

The geomagnetic conditions during this time were mild. The Disturbance storm time (Dst) index was
$>-30$ nT during the entirety of the observing period.\footnote{
    The Dst index was obtained from the World Data Center for Geomagnetism, Kyoto University
    (\url{http://swdcwww.kugi.kyoto-u.ac.jp/}). Accessed 2017-07-25.
}
Following the classification scheme of \citet{2008GMS...181.....K}, a weak geomagnetic storm has
$\text{Dst} < -30$ nT. Stronger geomagnetic storms have $\text{Dst} \ll -30$ nT.

Figure~\ref{fig:vtec} shows the median vertical total electron content (TEC) above OVRO measured
with the GIM \todo{citation}. The median is computed over all GPS measurements within 200 km of the
observatory. Over the observing period the TEC smoothly varies from 20 TECU at midday to 5 TECU
during the night. However, this measurement is sensitive to large scale fluctuations in the
ionosphere and doesn't capture small scale fluctuations.

@@

\begin{table*}[t]
    \centering
    \begin{tabular}{ccccccccc}
        \hline
        \hline
        & \tbf{Frequency} & \tbf{Bandwidth} & \tbf{FWHM} &
            \multicolumn2c{\tbf{Thermal Noise}} &&
            \multicolumn2c{\tbf{Singular Values}} \\
        \cline{5-6}
        \cline{8-9}
        \tbf{\#} & MHz & MHz & arcmin &
            K & Jy &&
            fraction $>\sqrt{\varepsilon}$ &
            fraction $>10\sqrt{\varepsilon}$ \\
        \hline
        1 & 36.528 & 0.024 &  &  & & & \\
        2 & 41.760 & 0.024 &  &  & & & \\
        3 & 46.992 & 0.024 &  &  & & & \\
        4 & 52.224 & 0.024 &  &  & & & \\
        5 & 57.456 & 0.024 &  &  & & & \\
        6 & 62.688 & 0.024 &  &  & & & \\
        7 & 67.920 & 0.024 &  &  & & & \\
        8 & 73.152 & 0.024 &  &  & & & \\
        \hline \hline
    \end{tabular}
    \caption{Summary of the generated all-sky maps\label{tab:summary}}
\end{table*}

@@

\begin{figure*}[p]
    \centering
    \begin{tabular}{c}
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw04} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw06} \\
    \end{tabular}
    \caption{
        channel maps r kewl
    }
    \label{fig:channel-maps}
\end{figure*}

\addtocounter{figure}{-1}
\begin{figure*}[p]
    \centering
    \begin{tabular}{c}
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw08} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw10} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw12} \\
    \end{tabular}
    \caption{
        continued
    }
\end{figure*}

\addtocounter{figure}{-1}
\begin{figure*}[p]
    \centering
    \begin{tabular}{c}
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw14} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw16} \\
        \includegraphics[height=0.32\textheight]{figures/channel-maps/spw18} \\
    \end{tabular}
    \caption{
        continued
    }
\end{figure*}

%\begin{figure*}[t]
%    \includegraphics[width=\textwidth]{figures/channel-maps/channel-maps}
%    \caption{
%        Mollweide-projected maps of the VHF sky north of $-30^\circ$ generated using the OVRO-LWA
%        and $m$-mode analysis imaging. The properties of each map are summarized in
%        Table~\ref{tab:summary}. The color scale is logarithmic and has units of Kelvin. As
%        described in the text, 8 bright sources have been subtracted. The small blank region near
%        the bottom of each map corresponds to the region of sky within 2 degrees of the Sun at the
%        time of the observations.
%    }
%    \label{fig:channel-maps}
%\end{figure*}
%
%\begin{figure*}[p]
%    \gridline{
%        \fig{figures/jackknife-maps/4.pdf}
%            {0.48\textwidth}{}
%        \fig{figures/jackknife-maps/6.pdf}
%            {0.48\textwidth}{}
%    }
%    \gridline{
%        \fig{figures/jackknife-maps/8.pdf}
%            {0.48\textwidth}{}
%        \fig{figures/jackknife-maps/10.pdf}
%            {0.48\textwidth}{}
%    }
%    \gridline{
%        \fig{figures/jackknife-maps/12.pdf}
%            {0.48\textwidth}{}
%        \fig{figures/jackknife-maps/14.pdf}
%            {0.48\textwidth}{}
%    }
%    \gridline{
%        \fig{figures/jackknife-maps/16.pdf}
%            {0.48\textwidth}{}
%        \fig{figures/jackknife-maps/18.pdf}
%            {0.48\textwidth}{}
%    }
%    \caption{
%        The jackknife estimated standard deviation corresponding to the maps presented in
%        Figure~\ref{fig:channel-maps}. The color scale is logarithmic and has units of Kelvin.
%    }
%    \label{fig:jackknife-maps}
%\end{figure*}
